---
title: "FinalProj"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
raw_data <- read.csv("train.csv")
```


## Data cleaning
```{r}
colSums(is.na(raw_data))
```

```{r}
data <- raw_data %>% replace_na(list(PoolQC = "None", 
                                     Fence = "None", 
                                     MiscFeature = "None"))
data <- data %>% replace_na(list(GarageFinish = "None",
                                     GarageQual = "None", 
                                     GarageCond = "None",
                                     GarageType = "None",
                                     GarageYrBlt = 0,
                                     FireplaceQu = "None"))
data <- data %>% replace_na(list(Electrical = "None",
                                     BsmtFinType2 = "None", 
                                     BsmtFinType1 = "None",
                                     BsmtExposure = "None",
                                     BsmtQual = "None",
                                     BsmtCond = "None"))
data <- data %>% replace_na(list(MasVnrArea = 0,
                                     MasVnrType = "None", 
                                     Alley = "None",
                                     LotFrontage = 0))
colSums(is.na(data))
```


## Data exploration 
```{r}
library(ggplot2)
library(ggcorrplot)
data_corr <- data %>% select(LotFrontage, LotArea, YearBuilt, 
                              YearRemodAdd, MasVnrArea, BsmtFinSF1,
                              BsmtFinSF2, BsmtUnfSF, TotalBsmtSF,
                              X1stFlrSF, X2ndFlrSF, LowQualFinSF,
                              GrLivArea, BsmtFullBath, BsmtHalfBath,
                              FullBath, HalfBath, BedroomAbvGr,
                              KitchenAbvGr, TotRmsAbvGrd, Fireplaces,
                              GarageYrBlt, GarageCars, GarageArea,
                              WoodDeckSF, OpenPorchSF, EnclosedPorch,
                              X3SsnPorch, ScreenPorch, PoolArea,
                              MiscVal, YrSold, MSSubClass, 
                              OverallQual, OverallCond, SalePrice)
correlations <- cor(data_corr)
ggcorrplot(round(correlations,2), lab = TRUE, tl.cex = 6, lab_size = 1, type = "lower")
```
Selected all continous variables and the three ordinal variables (MSSubClass, OverallQual and OverallCond) for the correlation plot. The last column SalePrice is the most important in the plot, we can see which variables are most correlated to the target and explain it the most. Overall Quality and total above ground square feet are the most correlated.



## Lasso
```{r}
library(glmnet)

data <- data %>% mutate_if(is.character,as.factor)
data$MSSubClass <- as.factor(data$MSSubClass)

#data <- data[-1299,]

train <- data[1:1100,]
train <- train[,-1]
test <- data[1101:1459,]
test <- test[,-1]

train$SalePrice <- log(train$SalePrice)
test$SalePrice <- log(test$SalePrice)

#rain_x <- model.matrix(train[,-c(1,81)])
train_x <- model.matrix(SalePrice ~ .,train)[,-81]
#test_x <- model.matrix(test[,-c(1,81)])
test_x <- model.matrix(SalePrice ~ .,test)[,-81]

lasso_cv <- cv.glmnet(train_x,train$SalePrice, alpha = 1)
plot(lasso_cv)

lasso_cv$lambda.min

lasso.pred=predict(lasso_cv,s=lasso_cv$lambda.min,newx=test_x)
#mean( (exp(lasso.pred)-exp(test$SalePrice))^2 ) 

#diff <- exp(lasso.pred)-exp(test$SalePrice)

diff <- data.frame(predictions = lasso.pred)
diff$residuals <- lasso.pred - test$SalePrice

diff$residab <- abs(diff$residuals)
diff$sq <- (diff$residuals * diff$residuals)

mean(diff$residab)

library(ggplot2)
ggplot(data = diff) + geom_point(mapping = aes(x=X1, y=residuals)) + 
  geom_abline(intercept = 0, slope = 0, colour = "red") + 
  ggtitle("Residual plot for Lasso Regression w/ outliers removed") + 
  xlab("predictions")


```

```{r}
model <- glmnet(train_x,train$SalePrice, alpha = 1, lambda = lasso_cv$lambda.min)
coef(model)
```

```{r}
train_temp <- model.matrix(~.,train)
train_temp <- as.data.frame(train_temp)
lm <- lm(SalePrice ~ .,data=train_temp)
summary(lm)
```

```{r}
test_temp <- model.matrix(~ .,test)
test_temp <- as.data.frame(test_temp)
test_temp <- test_temp[-199,]
library(caret)
predictions = predict(lm, newdata = test_temp)
```

```{r}
diff3 <- data.frame(predictions = predictions)
tempSalePrice <- test$SalePrice
tempSalePrice <- tempSalePrice[-199]
#diff3$predictions <- exp(diff3$predictions)
diff3$residuals <- predictions - tempSalePrice

diff3$residab <- abs(diff3$residuals)
diff3$sq <- (diff3$residuals * diff3$residuals)
#diff3 <- diff3 %>% mutate(sq = residuals^2)

mean(diff3$residab)


#diff3 <- diff3[-199,]




#mean(diff3$residab)
ggplot(data = diff3) + geom_point(mapping = aes(x=predictions, y=residuals)) + 
  geom_abline(intercept = 0, slope = 0, colour = "red") + 
  ggtitle("Residual plot for OLS Regression w/ outliers removed")
```


```{r}
par(mfrow=c(1,4))
plot(train$X1stFlrSF,train$SalePrice)
plot(train$OverallQual,train$SalePrice)
plot(train$GrLivArea,train$SalePrice)
plot(train$GarageArea,train$SalePrice)
```
```{r}
train2 <- data[0:1458,]
train_temp2 <- model.matrix(~.,train2)
train_temp2 <- as.data.frame(train_temp2)
lm2 <- lm(SalePrice ~ .,data=train_temp2)
predictions2 <- predict(lm2)
residuals2 <- predictions2 - train2$SalePrice
plot(predictions2,residuals2)
```

